---
title: ""
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(jsonlite)
library(httr)
library(dplyr)
library(sf)
library(leaflet)


library(EPIMOS)
```

# Introduction

The **EPIMOS** package (Environmental Prediction and Imputation Modular Statistical System) is a powerful tool designed to facilitate the analysis of environmental data. This package offers modular and flexible functionalities to access, process, and visualize location-based environmental measurements from various sources.

In this R Markdown document, we demonstrate how to use the EPIMOS package to fetch and process location data related to surface water monitoring. Our goal is to separate locations based on the availability of the "Q" parameter, which represents specific environmental conditions or measurements. Following this, we will visualize these locations on an interactive map.

## Workflow Overview

1. **Data Retrieval**:
    - We start by fetching location data from the OASI API using the `process_locations()` function. This function separates locations into two categories: those with the "Q" parameter and those without.

2. **Data Processing**:
    - Once the locations are categorized, we transform the coordinates from the Swiss coordinate system (CH1903) to the WGS84 system, which is commonly used in GPS and mapping applications.

3. **Data Visualization**:
    - Finally, we use the `leaflet` package to create an interactive map. This map allows users to visually explore the locations, with each point on the map representing a monitoring site. Clicking on these points reveals additional information about the site.

## Example Output

The processed data yields a list of locations that include the "Q" parameter, alongside those that do not. By converting the coordinates into the WGS84 system, we can accurately plot these locations on a global map, facilitating further exploration and analysis.

Below, we present an interactive map that visualizes these locations, offering an intuitive way to examine the geographical distribution of environmental monitoring sites.

```{r cars}

data <- process_locations("surfacewater")

# Convert to an sf object with Swiss coordinate system (CH1903)
data_sf <- st_as_sf(data, coords = c("coordinates.x", "coordinates.y"), crs = 21781)

# Transform to WGS84
data_wgs84 <- st_transform(data_sf, crs = 4326)

# Extract transformed coordinates
data_wgs84$longitude <- st_coordinates(data_wgs84)[,1]
data_wgs84$latitude <- st_coordinates(data_wgs84)[,2]

# Create a leaflet map
leaflet(data = data_wgs84) %>%
  addTiles() %>%
  addCircleMarkers(~longitude, ~latitude, popup = ~name, label = ~name, color = 'blue', radius = 5)
```


now training 


```{r}


library(tidyr)
library(ggplot2)

# Internal Function to process and append data
process_and_append_data <- function(response_data, parameter, location_name, data_frame) {
  data_lines <- strsplit(response_data, "\n")[[1]]
  data_lines <- data_lines[!grepl("^#", data_lines)]
  
  if (length(data_lines) > 0) {
    data_clean <- paste(data_lines, collapse = "\n")
    data_df <- read.csv(text = data_clean, sep = ";", header = TRUE)
    
    if ("data" %in% names(data_df) && parameter %in% names(data_df)) {
      data_df$data <- as.Date(data_df$data, format="%d.%m.%Y %H:%M")
      data_df[[parameter]] <- as.numeric(data_df[[parameter]])
      data_df$Location <- location_name
      data_frame <- rbind(data_frame, data_df)
    }
  }
  return(data_frame)
}

# Function to mask data for validation
mask_data_for_validation <- function(data_frame, prop_missing = 0.1) {
  set.seed(42)  # For reproducibility
  masked_data <- data_frame
  known_indices <- which(!is.na(masked_data), arr.ind = TRUE)
  n_to_mask <- round(prop_missing * nrow(known_indices))
  
  # Randomly select indices to mask
  mask_indices <- known_indices[sample(nrow(known_indices), n_to_mask), ]
  masked_data[mask_indices] <- NA
  
  return(list(masked_data = masked_data, original_indices = mask_indices))
}


```


